A Study of Common Bug Fix Patterns in Rust \thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}

### 2023-06-13 methodology.tex
more explicit

Figure~\ref{fig:extraction} illustrates path extraction on a simple change. The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of head to the callee instead of sending head itself. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non terminals. It is crucial to record the order of nodes in the path, so that we could record that order in our datapoint embedding. Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff. The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code. Instead, it is determined by the structural differences between the two ASTs of the code versions. The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.

@@

We propose a novel representation. Similar to the naive approach, we define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ collected from Syn. However, to account for order-sensitivity, we introduce two additional steps. Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each root-level scope. This captures the varying occurrence counts resulting from different element orders in the program. Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non-terminals that we consider more salient for Rust bugs. Section~\ref{sec:weighting_scheme}, immediately below,
illustrates this approach on an example.

@@

Multiplying the number of occurrences of all observed items in a change by their respective weights results in a vector of numbers. We call this vector the \emph{essence of change}. In this vector, the value for each item shows its importance in the change. Figure~\ref{fig:essence} shows the process of computing the essence of the code change that we saw in Figure~\ref{fig:extraction}. We have visualized this essence vector using a circle pack figure. The radius of each circle corresponds to the importance of the item inside it. The essence of change vector essentially serves as our embedding vector, transforming the patch illustrated in Figure~\ref{fig:extraction} into a fixed-sized vector as depicted in Figure~\ref{fig:essence}.

>>>>

\paragraph{Example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff. The change (shown in the top right corner of the Figure) replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non terminals. It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.

Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff. The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code. Instead, it is determined by the structural differences between the two ASTs of the code versions. The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing.

@@

We propose a novel representation. Similar to the naive approach, we define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ collected from Syn. However, to account for order-sensitivity, we introduce two additional steps. Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each root-level scope (e.g. structs, functions, impl blocks, etc). This captures the varying occurrence counts resulting from different element orders in the program. Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non-terminals that we consider more salient for Rust bugs.

\paragraph{Example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}. It also depicts the essence of change, to be described immediately below.

@@

Multiplying the number of occurrences of all observed items in a change by their respective weights results in a vector of numbers. We call this vector the \emph{essence of change}. In this vector, the value for each item shows its importance in the change.

\paragraph{Example: essence of change.} Figure~\ref{fig:essence} shows a complete example for computing the essence of change. The essence of change vector essentially serves as our embedding vector. The Figure shows the process for transforming the patch illustrated in Figure~\ref{fig:extraction} from fixed-size datapoints (computed as described in the previous subsection) into the essence of change vector. To better illustrate the essence vector, Figure~\ref{fig:essence} also graphically presents it using a circle pack figure. The radius of each circle corresponds to the importance of the item inside it.

### 2023-06-13 results.tex
more explicit

\noindent\textbf{Description:} This is a change that in all instances (in our benchmarks) has been proposed by Clippy---commit messages indicated the use of Clippy for this fix. The change applies on an object prior to calling \verb+to_string()+ on that object. In this change, a dereference operator is added to that object. The reason behind the change is that the object is a reference type of \verb+T+, while \verb+T+ directly implements \verb+to_string()+. Adding the dereference makes the compiler use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods\footnote{\url{https://rust-lang.github.io/rust-clippy/master/index.html\#inefficient\_to\_string}}. This change is aimed at improving performance, as the functionality remains unchanged.

>>>>

\noindent\textbf{Description:} This is a change that in all instances (in our benchmarks) has been proposed by Clippy---commit messages indicated the use of Clippy for this fix. The change applies on an object prior to calling \verb+to_string()+ on that object. In this change, a dereference operator is added to that object. The reason behind the change is that the object is a reference type of \verb+T+, where \verb+T+ directly implements \verb+to_string()+. Adding the dereference makes the compiler instead use the specialized implementation of \verb+to_string()+ and not go through slower string formatting methods\footnote{\url{https://rust-lang.github.io/rust-clippy/master/index.html\#inefficient\_to\_string}}. This change is aimed at improving performance; the functionality remains unchanged.
